---
title: "Selection for joint gain on negatively correlated traits"
author: "Seren Villwock"
date: "2023-07-14"
output: html_document
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(cache=FALSE)
knitr::opts_chunk$set(tidy=TRUE)
knitr::opts_chunk$set(results="hold")
knitr::opts_chunk$set(collapse=TRUE)
knitr::opts_chunk$set(fig.show="hold")
knitr::opts_chunk$set(tidy.opts = list(wrap=FALSE))

library(tidyverse)
library(AlphaSimR)
library(ggplot2)
set.seed(14850)
```


Generate a simulated population using `AlphaSimR`.

Simulation parameters are based on a cassava population with negatively correlated traits, dry matter and carotenoid content (color chart), estimated or directly from Rabbi et al. 2017. These parameters are:
DM & TC additive correlation = -0.2
DM & TC error correlation = -0.1
DM & TC means = 24.12, 6.53 (Parkes et al. 2020)
DM error variance = 41.08 (Parkes 2020 dominance + error variation)
TC error variance = 4.35
Number of founders = 200
Number of progeny = 6
HistoricalNe = 200
Segregating sites = 4400 (ICGMC 2015)
nQTL = 80
nSNP = 1000



```{r}
## Species, trait, and population parameters
DM_TC_adcorr = -0.2;
DM_TC_ercorr = -0.1;
DM_mean =  24.12;
DM_addvar = 18.44;
DM_errvar = 41.08;
TC_mean = 6.53;
TC_addvar = 1.36;
TC_errvar = 4.35;
FounderNum = 200;
numProgeny = 6

historicalNe <- 200
nFounders <- FounderNum
nChr <- 18
segSites <- 4400
nQTL <- 80
nSNP <- 1000
```

```{r, eval=F}
## Create founder population & generate traits

founderHaps <- runMacs2(nInd=nFounders, nChr=nChr, segSites=segSites,
                          Ne=historicalNe)

SP <- SimParam$new(founderHaps)

traitMeans <- c(DM=DM_mean, TC=TC_mean) 
#DM= dry matter; TC= total carotenoids/g fresh weight
#From Parkes et al. 2020

# Specify the variance and correlation
addVar <- c(DM=DM_addvar, TC=TC_addvar) 
#additive correlation matrix
addCor <- matrix(c(1, DM_TC_adcorr, DM_TC_adcorr, 1), nrow=2) 
# Specify the error correlation and calculate the covariance
errVar <- c(DM=DM_errvar, TC=TC_errvar)
errCor <- DM_TC_ercorr; errCov <- errCor*prod(sqrt(errVar))
errCov <- matrix(c(errVar[1], errCov, errCov, errVar[2]), nrow=2) #error covariance matrix
saveRDS(errCov, "./data/errCov.RDS")

SP$addTraitA(nQtlPerChr=nQTL, mean=traitMeans, var=addVar, corA=addCor)
SP$addSnpChip(nSnpPerChr=nSNP)

#save
saveRDS(SP, file="./data/SP.RDS")

# Create a founder population
founders <- newPop(founderHaps, simParam=SP)
nProgeny <- numProgeny
progenyPop <- randCross(founders, nCrosses=nFounders, nProgeny=nProgeny)
mtPhenos = setPheno(progenyPop,varE=errCov)

#Create a matrix with phenotypes and genotypes
simdata=cbind(progenyPop@id, mtPhenos@pheno, mtPhenos@gv)

founderdata <- list(mtPhenos, as.data.frame(simdata))
saveRDS(founderdata, file="./data/founderdata.RDS")
saveRDS(mtPhenos, file="./data/basepopulation.RDS")
```


Selection on residuals for simultaneous genetic gain 
```{r}
basePop = readRDS(file="./data/basepopulation.RDS")

popsize <- 200 
sim_length <- 50
nsim <- 100
run <- 1


#Plot the phenotypic and genetic mean through several generations of breeding.


allsims <- data.frame(meanGTrait1=NA, meanGTrait2=NA, sim=NA, generation=NA)

for(i in 1:nsim){

  # Store the results
  nGenerations = sim_length + 1 # +1 to store starting generation
  meanG = vector("list", length = nGenerations)
  varG = vector("list", length = nGenerations)
  
  # Save the starting values
  meanG[[1]] = meanG(basePop)
  varG[[1]] = varG(basePop)
  
  
  
  ## First selection step
  
  # Calculate composite trait to select on
  CalcStdRes <- function(Y){
    #Y is a matrix of trait values
    #fit a linear model
    trait_rel_mod <- lm(Y[,1] ~ Y[,2])
    
    #extract residuals
    return(trait_rel_mod$residuals)
  }
  
  CalcTraitRatio <- function(Y){
    #Y is a matrix of trait values
    ratio <- scale(Y[,1])/scale(Y[,2])
    return(ratio)
  }
  
  CalcStdResFlexible <- function(Y){
    #Y is a matrix of trait values
    
    #decide which trait to prioritize in this generation
    priority <- mean(scale(Y[,1])) > mean(scale(Y[,2]))
    
    #fit a linear model
    if(priority == T){
      trait_rel_mod <- lm(Y[,1] ~ Y[,2])} else{
        trait_rel_mod <- lm(Y[,2] ~ Y[,1])
      }
    
    #extract residuals
    return(trait_rel_mod$residuals)
  }
  
  CalcStdResIndex <- function(Y){
    #Y is a matrix of trait values
    
    #decide which trait to prioritize in this generation
    priority <- mean(scale(Y[,1])) > mean(scale(Y[,2]))
    
    #fit a linear model
    if(priority == T){
      trait_rel_mod <- lm(Y[,1] ~ Y[,2])} else{
        trait_rel_mod <- lm(Y[,2] ~ Y[,1])
      }
    
    #extract residuals
    resids <- trait_rel_mod$residuals
    
    #calculate index
    
    index <- 2*scale(resids) + 1*scale(Y[,1]) + 1*scale(Y[,2])
      
    return(index)
  }
  
  
  
  
  
  
  nSelected = 50
  newPopSelected = selectInd(pop = basePop,
                             nInd = nSelected,
                             use = "pheno",
                             trait = CalcStdResIndex,
                             simParam=SP)
  
  # Selection over many generations
  for (generation in 1:(nGenerations - 1)) {
    newPop = randCross(newPopSelected, nCrosses = nInd(basePop))
    newPop = setPheno(newPop, varE=errCov)
    newPopSelected = selectInd(pop = newPop,
                               nInd = nSelected,
                               use = "pheno",
                               trait = CalcStdResIndex,
                               simParam=SP)
    # Save summaries
    meanG[[1 + generation]] = meanG(newPop)
    varG[[1 + generation]] = varG(newPop)
    
  }
  
  # Save results
  meanGTrait1 = sapply(meanG, function(x) x[1])
  meanGTrait2 = sapply(meanG, function(x) x[2])
  meanRanges = range(c(meanGTrait1, meanGTrait2))
  
  varGTrait1 = sapply(varG, function(x) x[1, 1])
  varGTrait2 = sapply(varG, function(x) x[2, 2])
  varRanges = range(c(varGTrait1, varGTrait2))
  
  allsims <- rbind(allsims, cbind(meanGTrait1, meanGTrait2, sim=rep(run, length(meanGTrait2)), generation=1:nGenerations))
  
  print(i)
  run <- run + 1
                   
}
  

allsims_sum <- allsimsread %>% group_by(sim) %>%
  mutate(StdT1 = scale(meanGTrait1)) +
  mutate(StdT2 = scale(meanGTrait2))

summarise()
saveRDS(allsims, file="./simulationsrun100x.RDS")

ggplot(allsimsread, aes(generation, meanGTrait1)) +
  geom_point(color="black") +
  geom_smooth() +
  geom_point(aes(generation, meanGTrait2), color="orange") +
  xlab("Generation") +
  ylab("Scaled mean genetic value")


  # Plot mean of genetic values over time
  plot(x = 1:nGenerations, y = meanGTrait1, type = "l", col = "blue", lwd = 3,
       xlab = "Generation", ylab = "Mean of genetic values", ylim = meanRanges)
  lines(x = 1:nGenerations, y = meanGTrait2, type = "l", col = "blue", lty = 3, lwd = 3)
  legend(x = "topleft", legend = c("1", "2"), title = "Trait",
         lwd = 3, lty = c(1, 3), col = c("blue", "blue"))
  
  


```

